fsOGTT Pre- and Post-ETI

### 2025-08-12 Christine Chan/EnVision CF/fsOGTT_pre_post_ETI.qmd
converted c peptide units
manually fixed some ETI start dates per Monique
additional ETI date fixes

ids <- read_excel("./Data_Raw/ETI participant IDs_TV.xlsx")

@@

ids$`EnVision ID` <- sub("-2", "", ids$`EnVision ID`)

@@

@@

@@

@@

@@

# Plot

```{r}
df %>%
  group_by(study_id) %>%
  count(pre_post) %>%
  ggplot(., aes(y = study_id, x = n)) +
  geom_point() +
  theme_bw() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  facet_wrap(~pre_post)
```

- Adjust models for age
- Paneled figure with results for glucose, c-peptide, and insulin
- Look at acute insulin response (change in 30 minute c peptide)/(change in 30 minute glucose)
  - Can also do across the whole time period and can use insulin in the numerator also
- See Teams folder document for detailed plan

# Glucose

```{r}
# Make a long DF for models
long_df <- df %>% pivot_longer(Glucose_0:Glucose_180,
  names_to = c(".value", "timepoint"),
  names_pattern = "(.*)_(.*)",
  values_to = "Glucose"
)
long_df$timepoint <- factor(long_df$timepoint,
  levels = c("0", "10", "30", "60", "90", "120", "150", "180")
)
mod1 <- lmer(
  Glucose ~ timepoint * pre_post + age_visit + (1 | study_id),
  data = long_df
)
mod1 %>% tbl_regression(.)
means <- data.frame(emmeans(mod1, ~ timepoint + pre_post))
knitr::kable(means)
ggplot(means, aes(
  x = as.numeric(as.character(timepoint)), y = emmean,
  color = pre_post, group = pre_post
)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 1) +
  geom_line() +
  theme_bw() +
  xlab("Timepoint") +
  ylab("Glucose (mg/dL)") +
  scale_x_continuous(breaks = c(0, 10, 30, 60, 90, 120, 150, 180))
```

>>>>

ids <- read.csv("./Data_Clean/TV_ETI logitudinal patient list.csv", na.strings = "")
ids <- ids %>%
  select(EnVision.ID, GEM.ID) %>%
  rename(`EnVision ID` = EnVision.ID, `GEM ID` = GEM.ID)

@@

@@

# Manually correct a few per Monique
df$eti_start[df$study_id == "CC0001"] <- "2021-11-24"
df$eti_start[df$study_id == "CC0010"] <- "2023-01-16"
df$eti_start[df$study_id == "CC0022"] <- "2019-12-06"
df$eti_start[df$study_id == "CC0023"] <- "2020-01-09"
df$eti_start[df$study_id == "CC0026"] <- "2020-12-04"
df$eti_start[df$study_id == "CC0059"] <- "2021-06-23"

@@

df$pre_post[df$study_id == "CC0012"] <- "Pre"
df$pre_post[df$study_id == "CC0017"] <- "Pre"
df$pre_post[df$study_id == "CC0028"] <- "Pre"

@@

# Convert C-peptide level from pmol/L to ng/mL (multiply by 0.003)
df[, grep("C.Peptide_.*", colnames(df))] <-
  lapply(df[, grep("C.Peptide_.*", colnames(df))], function(c) {
    c * 0.003
  })

@@

# Calculate ODI, etc.

@@
